---
title: "Dairy ADPM Files"
format: html
execute:
  echo: false
  warning: false
  message: false
embed-resources: true
---

This is a document that explains the files use for ADPM dairy data
exercise. The data is from a real operation but has been processed to
make things a bit easier. The data was exported from DairyComp software.

-   The graphs are there to show the variable/column name

-   The table below the graphs contain the data and can be downloaded
    for your use.

```{r}
#| label: packages

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse,
  janitor,
  arrow,
  dtplyr,
  gt,
  lubridate,
  quarto, 
  zoo,
  tools,
  skimr,
  DT
  )

# load function
source("functions/fxn_dt_base.R")

# create
fxn_graph_file_skim<-function(data){
  
  sk_df <- skim(data)
  
  p<-ggplot(sk_df)+
  geom_bar(aes(x = skim_variable,
               y = complete_rate, fill = skim_type), stat = 'identity')+
  coord_flip()+
    theme_bw()+
    labs(x = "variable name",
         fill = "variable type")+
    scale_fill_viridis_d()
  
  print(p)
}

```

```{r}
#| label: import_parquet_files

# this imports all files at once 
files <- list.files('data/intermediate_files/',
                          pattern = "\\.parquet$", full.names = TRUE
                         ) 
names <- file_path_sans_ext(basename(files))

walk2(files, names, ~ assign(.y, read_parquet(.x), envir = .GlobalEnv))


```

```{r}
#| label: import_csv_files

# this imports all files at once 
# data source is everag dashboard
files_csv <- list.files('source_data/feed/',
                          pattern = "\\.csv$", full.names = TRUE
                         ) 
names_csv <- file_path_sans_ext(basename(files_csv))

walk2(files_csv, names_csv, ~ assign(.y, read_csv(.x, 
                                                  show_col_types = FALSE), envir = .GlobalEnv))


```

```{r}
#| label: parquet_data processing

# set dates
date_pull <- unique(animals$data_pull_date_max)
date_start <- date_pull-2000

animal_lact <- animal_lactations |> 
  filter(lact_number >0) |> 
  select(id_animal, id_animal_lact, date_fresh,
         lact_number, lact_group, date_dry) |> 
  left_join(animals |>  select(id_animal, breed, date_left),
            by = join_by(id_animal)
            )
  

# get last year events
events_1yr <- events_formatted |> 
    # remove repro events
  filter(event_type %in% c("health")) |> 
  filter(date_event>=date_start) |> 
  filter(lact_number >0) |> 
  select(id_animal, id_animal_lact, location_event, lact_number,
         lact_group_basic, date_archived, date_event, dim_event,
         date_concieved, date_aborted, date_repro_dx,
         event, event_type, protocols, remark) 

# sold deads
culls_1yr <- events_formatted |> 
    # remove repro events
  filter(event %in% c("SOLD", 'DIED')) |> 
  filter(date_event>=date_start) |> 
  filter(lact_number >0) |> 
  select(id_animal, id_animal_lact, lact_number,
         lact_group_basic, date_archived, date_event, dim_event,
         date_concieved, date_aborted, date_repro_dx,
         event, event_type, protocols, remark)


# get production data
production_1yr <- production_formatted |> 
    filter(date_test >= date_start) |>
  select(-c(source_file_path))
```

```{r}
#breeding data
# raw breeding data
# breeding data want 1st breeding age and conceived date
repro <- events_formatted |> 
  # select repro events
  filter(event %in% c("BRED")) |> 
  filter(lact_number >0) |> 
  select(id_animal, id_animal_lact, date_birth, breed,
         date_breeding = date_event,
         event, event_type, breeding_sire = remark, 
         breeding_result = R, 
         breeding_technician = T, breeding_code= B, 
         date_heat, date_concieved, date_aborted, date_repro_dx) |> 
  distinct() |> 
   group_by(id_animal_lact) |> 
  mutate(breeding_count = 1:n(), 
         breeding_total_count = sum(n()), 
         date_breeding_last = max(date_breeding),
         date_breeding_first = min(date_breeding)
         )|> 
  slice_max(order_by = breeding_count) |> 
  ungroup() |> 
  select(id_animal, id_animal_lact, date_birth, breed, date_breeding_last,
         breeding_sire, breeding_result, breeding_total_count,
         breeding_technician, breeding_code, 
         date_breeding_first,
         date_concieved, date_aborted, date_repro_dx
         ) |>
  # filter last so get right dates for first breeding etc.
  filter(date_breeding_last >= date_start) |> 
  # merge to get cow level info back
  left_join(animal_lact, by = join_by(id_animal, id_animal_lact, breed)) |> 
  select(id_animal, id_animal_lact, date_birth, date_fresh, 
         breed, lact_number, lact_group, everything())


# # PR tables
# breeding_1yr <- preg |> 
#   filter(period_start_date >= date_start) |> 
#   rename(date_period_start = period_start_date) |> 
#   select(-ends_with('risk'))

```


# Inventory

## Yearly

This table is a summary table that has counts of all the cows that were
at risk during the past year

```{r}

fxn_graph_file_skim(denominator_by_calendar_time_period)

```

```{r}

fxn_DT_base(denominator_by_calendar_time_period)

```

## by Lactation and DIM periods

```{r}

fxn_graph_file_skim(denominator_by_lact_group_time_period365)

```

```{r}

fxn_DT_base(denominator_by_lact_group_time_period365)

```

# Breeding Data files

These files have all the bred event info for the past year

## Raw breeding data.

```{r}

fxn_graph_file_skim(repro)

```

```{r}

fxn_DT_base(repro)

```


# Recorded Health Events

This has all the event data recorded for this dairy

```{r}

fxn_graph_file_skim(events_1yr)

```

```{r}

fxn_DT_base(events_1yr)

```

# Culling Events

This has all the culling data recorded for this dairy

```{r}

fxn_graph_file_skim(culls_1yr)

```

```{r}

fxn_DT_base(culls_1yr)

```

# DHI Data

This table is summary of the all the DHI test data that exists for this
dairy in the past year

```{r}

fxn_graph_file_skim(production_1yr)

```

```{r}

fxn_DT_base(production_1yr)

```

